{\large Gravitational wave sources in our Galactic backyard:}\\{\normalsize Predictions for BHBH, BHNS and NSNS binaries in LISA

### 2021-10-10 paper/tex/sections/variations.tex
update ce trend explanation

\subsection{Detection rates}\label{sec:detection_rate_analysis}
In Figure~\ref{fig:detection_rates}, we show the expected number of LISA detections for each model variation and discuss the prominent trends in the following sections. For the exact rates and uncertainties plotted in this figure see Table~\ref{tab:detection_rates}.

@@

Models \modAlphaLowest{}-\modAlphaHighest{} alter $\alpha_{\rm CE}$, the common-envelope efficiency parameter, to $0.1, 0.5, 2.0$ and $10.0$ respectively (where the fiducial model has $\alpha_{\rm CE} = 1.0$). Although we see that there are large variations in the LISA detection rates when changing $\alpha_{\rm CE}$, the detectable \textit{fraction} of systems remain largely unchanged. Therefore, any changes in the detection rate are due to changes in the \textit{formation} rate of the DCOs.

\todo{Explain formation rate differences with $\alpha_{\rm CE}$ changes. I'm pretty sure this comes down to higher alpha leading to larger separations but also fewer stellar mergers and this affects NSNSs and BHBH/BHNSs differently since the ejection of the envelope is very mass dependent. I'm making plots to show this (got sidetracked by orientation)}

>>>>

\subsection{Detection rates}\label{sec:detection_rate_analysis}
In Figure~\ref{fig:detection_rates}, we show the expected number of LISA detections for each model variation and discuss the prominent trends in the following sections. For the exact rates and uncertainties plotted in this figure see Table~\ref{tab:detection_rates} and for a plot of the rates relative to the fiducial model see Fig.~\ref{fig:dco_relative_rates}.

@@

Models \modAlphaLowest{}-\modAlphaHighest{} alter $\alpha_{\rm CE}$, the common-envelope efficiency parameter, to $0.1, 0.5, 2.0$ and $10.0$ respectively (where the fiducial model has $\alpha_{\rm CE} = 1.0$). Although we see that there are large variations in the LISA detection rates when changing $\alpha_{\rm CE}$, the detectable \textit{fraction} of systems remains largely unchanged. Therefore, any changes in the detection rate are due to changes in the \textit{formation} rate of the DCOs.

Altering $\alpha_{\rm CE}$ has two major effects on the formation rate of double compact objects. Recall that $\alpha_{\rm CE}$ represents the fraction of the orbital energy that is available for the ejection of the common-envelope. Therefore, increasing $\alpha_{\rm CE}$ allows common-envelope ejection to occur at larger separations (since more energy is available at any given separation). A second effect of increasing $\alpha_{\rm CE}$ is that common-envelopes are less likely to result in stellar mergers as more energy is available for envelope ejection. We can understand the trends in formation rate of the three DCO types by considering these two effects.

The BHBH rate peaks for $\alpha_{\rm CE} = 1$ (model \modFid{}) and reduces whether we increase \textit{or} decrease $\alpha_{\rm CE}$. This is because the increase in $\alpha_{\rm CE}$ shifts a significant fraction of the population to such large separations that they no longer merge within a Hubble time. Although the number of stellar mergers decreases as well, the overall change in the merging BHBH formation rate is negative. The inverse is also true for decreasing $\alpha_{\rm CE}$, where BHBHs have shorter separations but stellar mergers are more frequent. Thus we find that $\alpha_{\rm CE} = 1$ seems to be optimal for BHBHs by balancing these two effects most effectively.

The BHNS rate follows the same pattern as BHBHs. This is expected since the BH is formed first for every BHNS in our sample and so the common-envelope event is due to a RLOF of BH progenitor, just as with BHBHs. However, given the similarity in the BHNS rates between $\alpha_{\rm CE} = 1$ and $\alpha_{\rm CE} = 2$ (models \modFid{} and \modAlphaHigh{}), the `optimal' $\alpha_{\rm CE}$ for BHNSs seems to be somewhere between 1 and 2 rather than very close to 1 as with BHBHs.

In contrast, $\alpha_{\rm CE} = 1$ is far from optimal for NSNSs. Increasing $\alpha_{\rm CE}$ (models \modAlphaHigh{} and \modAlphaHighest{}) results in significantly higher rates. This is because LISA detectable NSNSs tend to have been formed very recently before detection (see Fig.~\ref{fig:fiducial_pdf_distributions}e). Therefore, though higher $\alpha_{\rm CE}$ results in larger \textit{initial} separations, this does not adversely affect the NSNS rate as the systems are simply detected later after their formation compared to the fiducial model. So for models \modAlphaHigh{} and \modAlphaHighest{}, fewer systems result in stellar mergers and the overall population remains within the LISA band, producing much higher detection rates. %Finally, for model \modAlphaLowest{}, the reduction in common-envelope efficiency is strong enough that a sufficient number of systems are formed at shorter separations to offset those lost to stellar mergers.
